Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-31 Paper.Rmd
modified

Similarly, we applied multivariate analysis on four spatial datasets; \textbf{(i)} aphid-count and \textbf{(ii)} aphid-firstflight date were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species. We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, for aphid-abundance data and for aphid-firstflight data, only 10 locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}) were considered. \textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013). These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).--> [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}). \textbf{(iv)} Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.

>>>>

Similarly, we applied multivariate analysis on four spatial datasets; \textbf{(i)} aphid-count and \textbf{(ii)} aphid-firstflight date were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species (Table \ref{SM-tab_aphid_info}). We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, for aphid-abundance data and for aphid-firstflight data, only 10 locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}) were considered. \textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species (Table \ref{SM-tab_plankton_info}) and for 56 years (1958-2013). These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).--> [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}). \textbf{(iv)} Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.

### 2018-10-31 SupportingInformation.Rmd
modified

@@

>>>>

```{r tab_aphid_info, echo=F, results='asis',message=F}
library(tibble)
aphid_info <- tibble(c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
           "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
           "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach potato aphid",
           "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
           c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
             "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",
             "Brachycaudus helichrysi","Brevicoryne brassicae","Hyalopterus pruni","Acyrthosiphon pisum","Myzus persicae",
             "Macrosiphum euphorbiae","Metopolophium dirhodum","Myzus ascalonicus","Drepanosiphum platanoidis","Cavariella aegopodii")
)
knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 20 aphids commonly found across UK \\label{tab_aphid_info}",
      col.names = c("Common name", "Latin binomial"))
```

@@

```{r tab_plankton_info, echo=F, results='asis',message=F}
library(tibble)
plankton_info <- tibble(c('Calanus I-IV','Para-Pseudocalanus spp','Temora longicornis','Acartia spp',
                          'Centropages typicus',
                          'Oithona spp','Echinoderm larvae','Calanus finmarchicus',
                          'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
                   'Euphausiacea','Thalassiosira spp','Rhizosolenia styliformis',
                   'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
                   'Rhizosolenia alata alata','Pseudocalanus elongatus Adult',
                   'Nitzschia delicatissima','Nitzschia seriata')
)
knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 22 planktons found across UK seas\\label{tab_plankton_info}",
      col.names = c("Plankton sp. name"))
```
